Validity of a model for a cantilever two-dimensional beam

### 2024-04-06 article_old.tex

% \documentclass[11pt]{elsarticle}
\input{preamble_article.tex}
\usepackage{geometry}
\geometry{
  left=1.3in,
  right=1.4in,
  top=1.4in,
  bottom=1.4in,
}

@@

\title{Comparison of linear beam theories}

\author{R. du Toit\corref{cor1}}
\ead{u15000258@tuks.co.za}
\cortext[cor1]{Corresponding author}
\author{N.F.J. van Rensburg}
\address{Department of Mathematics, University of Pretoria, Pretoria, South Africa}

\journal{Applied Mathematics and Computation}

\begin{abstract}
Beams, while commonly used in engineering, are complex three-dimensional objects. Simplified models are often used in vibration problems due to their reduced complexity and computational demand. However, the accuracy of these models in practical applications is not always certain. This article extends previous research to examine a two-dimensional cantilever beam model's validity using modal analysis and the Finite Element Method (FEM), comparing its solutions with those of a three-dimensional model. Our findings reveal that the beam's shape is a crucial factor in the model's validity for practical applications. For the appropriate beam shapes, the two-dimensional model is a valid approximation of the three-dimensional model. 
\end{abstract}

\begin{keyword}
	Beam \sep Cantilever Beam \sep Two-Dimensional Model \sep Three-Dimensional Model \sep Vibration Analysis \sep Modal Analysis \sep Finite Element Method
\end{keyword}

@@

\section{Introduction}
In the article \cite{LVV09}, the authors investigate the validity of a cantilever Timoshenko beam model. Validity refers to how well the solution of the simplified model compare to a more complex higher dimensional model. The reference model in \cite{LVV09} is a two-dimensional cantilever beam model.

In this article, we build upon \cite{LVV09} by further examining the validity of a two-dimensional model for a cantilever beam, using a three-dimensional model as the reference. We follow the same methodology as \cite{LVV09}, employing modal analysis and the Finite Element Method (FEM) to compare the two models. This extention is motivated by the authors of \cite{LVV09}.

To start, the models are presented in a dimensionless variational form. We explain how the two-dimensional model is derived from the three-dimensional model as a special case of plane stress. The well-posedness of the two models is not proved in this article, but the necessary existence and uniqueness results from \cite{VV02} are mentioned.

Next we introduce the concept of modal analysis with a general vibration problem. Through modal analysis, we derive a formal series solution for the vibration issue, which is a linear combination of the problem's eigenvectors and eigenvalues. This approach allows us to compare the solutions of the two models by focusing on a comparison of their eigenvectors and eigenvalues.

We then apply the Finite Element Method (FEM) to the two models and obtain a formula for the models to numerically solve the eigenvalue problem. The article \cite{BV13} provides essential convergence results for the FEM in vibration problems. The textbook \cite{SF73} also offers insights into the convergence of eigenvalues and eigenfunctions in a vibration problem when using FEM.

Following \cite{LVV09}, by considering the mode shapes, we can match up the eigenvalues and compare them. In the comparison, we consider different shapes of beams that are useful in practical applications. We find that the shape of the beam significantly influences the validity of the two-dimensional model. The results of the comparison are presented in the form of tables and graphs.

The layout of the paper is as follows:
\begin{itemize}
\item Section \ref{sec:LinearMathematicalModels} introduces the linear mathematical models for the three-dimensional, two-dimensional, and Timoshenko beam models.
\item Section \ref{sec:ModalAnalysis} covers modal analysis, compares linear models, and touches on the models' well-posedness.
\item Section \ref{ssec:2D_Model:FEM} outlines the variational problem for the two-dimensional model and applies the Finite Element Method to derive an eigenvalue problem.
\item Section \ref{ssec:3D_Model:FEM} outlines the variational problem for the three-dimensional model and applies the Finite Element Method to derive an eigenvalue problem.
\item Section \ref{sec:validity-of-a-cantilever-timoshenko-beam} examines the validity of the Timoshenko model for a cantilever beam, drawing on results from \cite{LVV09}.
\item Section \ref{sec:validity-of-a-2d-beam} expands on \cite{LVV09} to assess the validity of a two-dimensional model for a cantilever beam.
\item Section \ref{sec:conclusion} concludes the paper.
\end{itemize}

\section{Linear mathematical models} \label{sec:LinearMathematicalModels}
	\subsection{Three-dimensional theory}
		Consider a vector valued displacement function $u$ defined on $\Omega \subset R^3$.

@@

		\subsubsection*{Constitutive Equation}
			For an isotropic material, Hooke's law can be written in terms of Young's modulus $E$ and Poisson's ratio $\nu$. And if the principal stresses $\sigma_i$ are all non-zero, then Hooke's law can be written in the following form:
			\begin{eqnarray}
				T = \left( \frac{E}{1+\nu} \right)\mathcal{E} + \frac{\nu E}{(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \label{eq:3D_Model:CE}.
			\end{eqnarray} Here $\mathcal{E}$ is the strain tensor and $T$ is the stress tensor.

@@

			Let $\ell$ represent a significant dimension (e.g., length) of an elastic body and $G$ the shear modulus. Define dimensionless variables and normalized functions as
			\[\tau = \frac{t}{t_0},\ \xi_i = \frac{x_i}{\ell},\ u^*(\xi,\tau) = \frac{u(x,t)}{\ell},\ Q^{*} = \ell G \kappa^2 Q,\ \sigma_{ij}^*(\xi,\tau) = \frac{\sigma_{ij}(x,t)}{G\kappa^2},\]
			where $\kappa^2$ is a dimensionless constant and $\displaystyle t_0 = \ell\sqrt{\frac{\rho}{G\kappa^2}}$. Let $\displaystyle \gamma= \frac{G\kappa^2}{E}$.

			Substituting these into the equation of motion \eqref{eq:3D_Model:EM} gives
			\[\partial_{\tau} u^{*} = \text{div}T^* + Q^*,\]
			where $\displaystyle T^* = \sigma_{i,j}^*(\xi,\tau)$. The dimensionless forms for \eqref{eq:3D_Model:HL} and \eqref{eq:3D_Model:CE} are
			\[\mathcal{E} = \gamma(1+\nu)T^* - \gamma\nu \text{Tr}(T^*)I,\]
			\[T^* = \frac{\mathcal{E}}{\gamma(1+\nu)} + \frac{\nu\text{Tr}(\mathcal{E})I}{\gamma(1+\nu)(1-2\nu)}.\]

			The choice of dimensionless constants follow \cite{LVV09} and allows for a comparison between this model and the Timoshenko beam model.

		\textbf{Equations of motion in dimensionless form}
			\begin{eqnarray}
				\partial_t^2 u & = & \textrm{div}T + Q \label{eq:3D_Model:EM-D}
			\end{eqnarray}
			with
			\begin{eqnarray}
				\textrm{div}  T & = &
				\begin{bmatrix}
					\partial_1 \sigma_{11} + \partial_2 \sigma_{12} + \partial_3 \sigma_{13} \\
					\partial_1 \sigma_{21} + \partial_2 \sigma_{22} + \partial_3 \sigma_{23} \\
					\partial_1 \sigma_{31} + \partial_2 \sigma_{32} + \partial_3 \sigma_{33}
				\end{bmatrix}.\label{eq:3D_Model:divT-D}
			\end{eqnarray}

		\textbf{Constitutive equations in dimensionless form}
			\begin{eqnarray}
				T = \frac{1}{\gamma(1+\nu)} \mathcal{E} + \frac{\nu}{\gamma(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \label{eq:3D_Model:CE-D}
			\end{eqnarray}

		\subsubsection*{Cantilever model}
			Find a vector valued function $u$, satisfying equations
			\eqref{eq:3D_Model:EM-D} to \eqref{eq:3D_Model:CE-D} and the boundary
			conditions $\displaystyle u = 0$ on part of the boundary and $\displaystyle Tn = 0$ on the rest of the boundary. This will be referred to as \textbf{Problem 3D}.

	\subsection{Two-dimensional theory}
		\subsubsection*{General plane stress}
			Stresses acting parallel to a single plane are referred to as plane stress. 
			Consider the right-hand orthonormal set $\left\{e_1, e_2, e_3\right\}$. Without
			loss of generality, assume the stresses act parallel to the $e_1$-$e_2$ plane
			i.e. $\sigma_{3i} = 0$ for all $i$. From Hooke's Law \eqref{DM_H_E} the
			following equations are obtained: \label{sym:e_i}

@@

			\begin{minipage}{.5\linewidth}
				\begin{eqnarray*}
					\sigma_{11} & = & \frac{1}{\gamma(1-\nu^2)} ( \varepsilon_{11} + \nu\varepsilon_{22})\\
					\sigma_{22} & = & \frac{1}{\gamma(1-\nu^2)} (\nu \varepsilon_{11} + \varepsilon_{22}).
				\end{eqnarray*}
			\end{minipage}%
			\begin{minipage}{.5\linewidth}
				\begin{eqnarray}
					\sigma_{12} & = & \frac{1}{\gamma(1+\nu)} \varepsilon_{12} \label{stress_comp}
				\end{eqnarray}
			\end{minipage}

@@

			This is known as general plane stress. The two-dimensional model is a special case of plane stress.

		\subsubsection*{The two dimensional model}
			Assume that $\sigma_{3i} = 0$ for $i = 1, 2, 3$; $\partial_3 u_1 = 0$, $\partial_3 u_2 = 0$ ($u_1$ and $u_2$ are functions of $x_1$ and $x_2$), and the strain component $\varepsilon_{33} = 0$. Then using the definition of strain 
			\begin{eqnarray}
				\partial_i u_3 = 0 \quad \textrm{ for } i = 1,2,3 \textrm{ on } \Omega.
			\end{eqnarray}

@@

			The equations of motion and constitutive equations follow:
			\subsubsection*{Equation of motion}\label{sssec:2D_Model:EquationOfMotion}
			\begin{eqnarray}
				\partial_t^2 u & = & \textrm{div}T + Q, \label{eq:2D_Model:EM}
			\end{eqnarray}
			where
			\begin{eqnarray}
				\textrm{div} T & = &
				\begin{bmatrix}
					\partial_1 \sigma_{11} + \partial_2 \sigma_{12} \\
					\partial_1 \sigma_{21} + \partial_2 \sigma_{22}
				\end{bmatrix}. \label{eq:2D_Model:DivT}
			\end{eqnarray}
			\subsubsection*{Constitutive equation}\label{sssec:2D_Model:ConstitutiveEquation}
			\begin{eqnarray}
				T = \frac{1}{\gamma(1+\nu)}\mathcal{E} + \frac{\nu}{\gamma(1-\nu^2)}\textrm{Tr}(\mathcal{E}) I. \label{eq:2D_Model:CE}
			\end{eqnarray}

		\subsubsection*{Cantilever model}

			Find a vector valued function $u$, satisfying equations \eqref{eq:2D_Model:EM} to \eqref{eq:2D_Model:CE} and the boundary conditions $u = 0$ on part of the boundary and $Tn = 0$ on the rest of the boundary. This will be referred to as \textbf{Problem 2D}.

	\subsection{Timoshenko theory}
		Consider a Timoshenko beam with no damping defined on the interval $[0,\ell]$.
		Let $w$ represent the transverse displacement and $\phi$ a rotation of the cross-sections of the beam.

		\subsubsection*{Equations of motion}\label{sssec:1D_Model:EquationOfMotion}
			\begin{eqnarray}
				\rho A \partial_t^2 w &=& \partial_x V + Q, \label{eq:1D_Model:EquationOfMotion1}\\
				\rho I\partial_t^2 \phi  &=& V + \partial_x M. \label{eq:1D_Model:EquationOfMotion2}
			\end{eqnarray}\label{sym:A}\label{sym:V}\label{sym:Iinertia}\label{sym:M}\label{sym:phiT}
			In~\eqref{eq:1D_Model:EquationOfMotion1} and \eqref{eq:1D_Model:EquationOfMotion2} $\rho$ denotes the density, $A$ is the area of a cross section, $I$ is the area moment of inertia, $M$ is the moment, $V$ is the shear force and $Q$ is an external force acting on the beam.

		\subsubsection*{Constitutive equations}\label{sssec:1D_Model:ConstitutiveEquation}
			\begin{eqnarray}
				M &=& EI\partial_x \phi, \label{eq:1D_Model:ConstitutiveEquations1}\\
				V &=& AG\kappa^2(\partial_x w - \phi), \label{eq:1D_Model:ConstitutiveEquations2}\label{sym:kappa2T}
			\end{eqnarray}
			$E$ is Young's modulus, $G$ the shear modulus and $\kappa^2$ the shear correction factor.

		\subsubsection*{Dimensionless form}\label{sssec:1D_Model:DimensionlessForm}
			As mentioned in the dimensionless form of the three-dimensional model, the same dimensionless scaling can be applied to the Timoshenko beam theory.

			Set \[\tau = \frac{t}{t_0},\ \xi = \frac{x}{\ell},\ w^*(\xi,\tau) = \frac{w(x,t)}{\ell},\ \phi^*(\xi, \tau) = \phi(x,t).\]

			Dimensionless shear force, moment, and external force density are \[V^{*} = \frac{V}{AG\kappa^2},\ M^{*} = \frac{M}{A G\kappa^2 \ell},\ Q^* = \frac{Q\ell}{A G\kappa^2}.\]

			Use $\displaystyle t_0 = \ell\sqrt{\frac{\rho}{G\kappa^2}}$ as before.

			Using the dimensionless constants from \cite{LVV09}, 
			\[\alpha = \frac{A \ell^2}{I},\ \beta =\frac{AG\kappa^2 \ell^2}{EI},\]
			with \[\frac{\beta}{\alpha} = \gamma,\] where $\gamma$ is the dimensionless parameter from the three-dimensional model.

		\subsubsection*{Dimensionless equations of motion}\label{sssec:1D_Model:DimensionlessEquationsOfMotion}
			\begin{eqnarray}
				\partial_{t}^{2} w &=& \partial_{x}V + Q, \label{eq:1D_Model:EquationOfMotion1D}\\
				\frac{1}{\alpha} \partial_{t}^{2} \phi &=& V + \partial_{x}M.\label{eq:1D_Model:EquationOfMotion2D}
			\end{eqnarray}
		\subsubsection*{Dimensionless constitutive equations}\label{sssec:1D_Model:DimensionlessConstitutiveEquations}
			\begin{eqnarray}
				M &=& \frac{1}{\beta}\partial_x \phi, \label{eq:1D_Model:ConstitutiveEquations1D}\\
				V &=& \partial_x w-\phi. \label{eq:1D_Model:ConstitutiveEquations2D}
			\end{eqnarray}
		The original notation is retained for convenience.

		\textbf{Models and variational forms}
			\subsubsection*{Bilinear forms and integral}
				For $f,g \in T[0,1]$, define the bilinear forms
				\begin{eqnarray}
					c(f,g) & = & \int_0^1  \partial_t^2 f_1 g_1 + \frac{1}{\alpha}\int_0^1  \partial_t^2 f_2g_2, \label{eq:1D_Model:Bilinear} \\
					b(f,g) & = &   \int_0^1 (f_1'-f_2)(g_1' - g_2) + \frac{1}{\beta}\int_0^1 f_2' g_2', \label{eq:1D_Model:Bilinear_c}
				\end{eqnarray}
				Define the integral
				\begin{eqnarray}
				(f,g)
					&=& \int_{0}^1 fg \label{eq:1D_Model:Bilinear_int}
				\end{eqnarray} for all $f,g \in L^2(0,1).$

				\textbf{Remark}
				$L^2(a,b)$ is the space of all square integrable functions on the interval $(a,b)$. The inner product is defined by $\displaystyle (f,g) = \int_a^b fg$, and the induced norm $\displaystyle ||f|| = \int_a^b f^2$.\\

			\subsection*{Pinned-pinned beam}
				Find functions $w$ and $\phi$ satisfying equations \eqref{eq:1D_Model:EquationOfMotion1D} to \eqref{eq:1D_Model:ConstitutiveEquations2D} and the boundary conditions:

				{Boundary Conditions}
				\begin{eqnarray*}
					w(0,\cdot) = 0, \ \ &M(0,\cdot) = 0, \label{eq:1D_Model:ProblemT1BC1}\\
					w(1,\cdot) = 0, \ \ &M(1,\cdot) = 0. \label{eq:1D_Model:ProblemT1BC2}
				\end{eqnarray*}

				This will be referred to as \textbf{Problem T-1}. Define the set of test functions $T[0,1]$ for Problem T-1 as
				\begin{eqnarray*}
					T[0,1] &:=&  \left\{f \in C^1[0,1] \ | \ f(0) = f(1) = 0 \right\} \times C^1[0,1]
				\end{eqnarray*}

				Then the variational problem for Problem T-1 is given by
				\subsubsection*{Problem T-1V}\label{sssec:1D_Model:ProblemT1V}
					Find a function ${u} = \langle w, \phi \rangle$ such that for all $t >0$, ${u} \in  T[0,1]$ satisfying
					\begin{eqnarray}
						c(\partial_t^2 u,{\phi}) &=& -b({u},{\phi}) + (Q,{\phi})
					\end{eqnarray} for each ${\phi} = \langle v, \psi \rangle \in T[0,1]$.

			\subsection*{Cantilever beam}
				Find functions $w$ and $\phi$ satisfying equations \eqref{eq:1D_Model:EquationOfMotion1D} to \eqref{eq:1D_Model:ConstitutiveEquations2D} and the boundary conditions:

				{Boundary Conditions}
				\begin{eqnarray*}
					w(0,\cdot) = 0, \ \ &\phi(0,\cdot) = 0, \label{eq:1D_Model:ProblemT2BC1}\\
					M(1,\cdot) = 0, \ \ &V(1,\cdot) = 0. \label{eq:1D_Model:ProblemT2BC2}
				\end{eqnarray*}

				This will be referred to as \textbf{Problem T-2}.  Define the set of test functions $T[0,1]$ for Problem T-2 as
				\begin{eqnarray*}
					T[0,1] &:=& \left\{g \in C^1[0,1] \ | \ g(0) = 0 \right\} \times \left\{g \in C^1[0,1] \ | \ g(0) = 0 \right\}
				\end{eqnarray*}

				Then the variational problem for Problem T-2 is given by
				\subsubsection*{Problem T-2V}\label{sssec:1D_Model:ProblemT1V}
					Find a function ${u} = \langle w, \phi \rangle$ such that for all $t >0$, ${u} \in  T[0,1]$ satisfying
					\begin{eqnarray}
						c(u,{\phi}) &=& -b({u},{\phi}) + (Q,{\phi}) \label{var_form_timo}
					\end{eqnarray} for each ${\phi} = \langle v, \psi \rangle \in T[0,1]$.

\section{Modal analysis and comparison of linear models}\label{sec:ModalAnalysis}

	\subsection*{Existance and uniqueness of solutions}
		Our two models, the three-dimensional modela and the two-dimensional model, concern the vibration of elastic bodies. These type of models have a similar variational form as the wave equation and are called second order hyperbolic type problems. The articles \cite{VV02} and \cite{VS18} provide existence and uniqueness results for these type of problems. In \cite{VV02}, it is assumed that $b$ is symmetric. \cite{VS18} extends the results of \cite{VV02} to the case where $b$ is not symmetric. For the models in the article, \cite{VV02} is a sufficient reference.

@@

		Consider a general second order hyperbolic type problem GVar. This general problem is used by the 
		authors of \cite{VV02} to prove existence and uniqueness results. The models in this article are 
		special cases of this general problem.

>>>>

\documentclass[preprint,12pt]{elsarticle}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{draftwatermark}
\usepackage{tabularx} % Required for tables that stretch to text width
\usepackage{booktabs} % For prettier tables
\usepackage{ltablex}
\usepackage{xcolor,cancel,bm,tikz}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{amsmath,amssymb,amsthm,amsfonts}
\usepackage{enumitem}
\usepackage{kantlipsum,graphicx}
\usepackage{multicol,colortbl,multirow}
\usepackage{pgfplots,booktabs,xfrac,floatrow}
\usepackage{color}
\usepackage{titlesec}
\usepackage{natbib}

\titleformat*{\section}{\Large\bfseries}
\titleformat*{\subsection}{\normalsize\bfseries}
\titleformat*{\subsubsection}{\normalsize\bfseries}
\numberwithin{equation}{section}

\usepackage{setspace}
\setstretch{0.7} % Adjust this value as needed

\setlength{\parindent}{0pt}

\usepackage{geometry}
\geometry{
  left=1in,
  right=1in,
  top=1in,
  bottom=1in,
}

\SetWatermarkText{DRAFT}
\SetWatermarkScale{1}

\journal{Mathematical and Computer Modelling}

@@

\title{Validity of a model for a cantilever two-dimensional beam}

\author{Rudi du Toit}
\address{University of Pretoria, Pretoria, South Africa}

\begin{abstract}
Your abstract here.
\end{abstract}

\begin{keyword}
Vibration \sep Cantilever \sep Beam \sep Three-Dimensional \sep Two-Dimensional
\end{keyword}

@@

\section{Introduction}
Your introduction here.

\section{Models}

\subsection{Three-dimensional model}
Consider a vector valued displacement function $u \in \Omega \subset R^3$.

@@

\subsubsection*{Constitutive Equation}
For an isotropic material, Hooke's law can be written in terms of Young's modulus $E$ and Poisson's ratio $\nu$ as follows.\\

\textit{Hooke's law}
\begin{eqnarray}
	\mathcal{E} = \left( \frac{1+\nu}{E} \right)T - \frac{\nu}{E}\textrm{Tr}(T)I,  \label{eq:3D_Model:HL}
\end{eqnarray} The components of the strain tensor $\mathcal{E}$ are defined as
\begin{eqnarray}
	\varepsilon_{ij} = \frac{1}{2}\left( \frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i} \right). \label{eq:3D_Model:Strain}
\end{eqnarray}\label{sym:varepsilonij}

If the principal stresses $\sigma_i$ are all non-zero, then Hooke's law can be written in an alternative form.\\

\textit{Hooke's law in the alternative form}
\begin{eqnarray}
	T = \left( \frac{E}{1+\nu} \right)\mathcal{E} + \frac{\nu E}{(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \label{eq:3D_Model:CE}.
\end{eqnarray}
With $\textrm{Tr}(\mathcal{E}) = \epsilon_{11} + \epsilon_{12} + \epsilon_{13}$ is the trace of the strain operator $\mathcal{E}$.\\

@@

\subsection{Dimensionless form} \label{ssec:3D_Model:DimensionlessForm}
Suppose $\ell$ represents some notable dimension (e.g.the length) of the elastic body 
and $G$ the shear modulus of elasticity.\\

Set \[\tau = \frac{t}{t_0},\quad \xi_i = \frac{x_i}{\ell},\quad {u}^*(\xi,\tau) = \frac{1}{\ell}{u}(x,t), \quad Q^{*} = \ell G \kappa^2 Q, \text{and} \ \,\,\, \sigma_{ij}^*(\xi,\tau) = \frac{1}{G\kappa^2}\sigma_{ij}(x,t),\]\label{sym:t}\label{sym:G}\label{sym:tau}\label{sym:xi}\label{sym:Qstar}\label{sym:sigmaijstar}\label{sym:ell}\label{sym:kappa2}

where $\kappa^2$ is a dimensionless constant and $t_0$ must be specified. Choose \[t_0 = \ell\sqrt{\frac{\rho}{G\kappa^2}} \textrm{ and } \gamma= \frac{G\kappa^2}{E}.\]

Substitution into \eqref{eq:3D_Model:EM} and yields
\begin{eqnarray*}
	\partial_{\tau} u^{*} = \textrm{div}T^* + Q^* \textrm{ where } T^* = \sigma_{i,j}^*(\xi,\tau).
\end{eqnarray*}

The dimensionless forms of \eqref{eq:3D_Model:HL} and \eqref{eq:3D_Model:CE} are
\begin{eqnarray*}
	\mathcal{E} = \frac{G \kappa^2}{E} \left[ (1+\nu)T^* - \nu \textrm{Tr}(T^*)I \right],
\end{eqnarray*}\label{sym:I}

and
\begin{eqnarray*}
	T^* = \frac{E}{G \kappa^2} \left[\left( \frac{1}{1+\nu} \right)\mathcal{E} + \frac{\nu}{(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \right].
\end{eqnarray*}

Using the dimensionless constant $\gamma$,
\begin{eqnarray}
	\mathcal{E} = \gamma(1+\nu)T^* - \gamma\nu \textrm{Tr}(T^*)I \label{DM_H_E}
\end{eqnarray}
and
\begin{eqnarray}
	T^* = \frac{1}{\gamma(1+\nu)}\mathcal{E} + \frac{\nu}{\gamma(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I. \label{DM_H_T}
\end{eqnarray}

\textit{Remark} The constants introduced are the same as those in \cite{LVV09}. They simplify the equations and allow for comparisons to other models such as the Timoshenko beam model, although that is not covered in this article.\\

The original notation is retained for convenience.

\subsubsection*{Equations of motion in dimensionless form}\label{sssec:3D_Model:EquationOfMotionDimensionless}
\begin{eqnarray}
	\partial_t^2 u & = & \textrm{div}T + Q \label{eq:3D_Model:EM-D}
\end{eqnarray}
with
\begin{eqnarray}
	\textrm{div}  T & = &
	\begin{bmatrix}
		\partial_1 \sigma_{11} + \partial_2 \sigma_{12} + \partial_3 \sigma_{13} \\
		\partial_1 \sigma_{21} + \partial_2 \sigma_{22} + \partial_3 \sigma_{23} \\
		\partial_1 \sigma_{31} + \partial_2 \sigma_{32} + \partial_3 \sigma_{33}
	\end{bmatrix}.\label{eq:3D_Model:divT-D}
\end{eqnarray}

\subsubsection*{Constitutive equations in dimensionless form}\label{sssec:3D_Model:ConstitutiveEquationsDimensionless}
\begin{eqnarray}
	T = \frac{1}{\gamma(1+\nu)} \mathcal{E} + \frac{\nu}{\gamma(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \label{eq:3D_Model:CE-D}
\end{eqnarray}

\subsection{Model problems}\label{ssec:3D_Model:ModelProblems}
Divide the boundary of $\Omega$ into two distinct parts, referred to as $\Sigma$ and $\Gamma$. The following will be considered a model problem for a three-dimensional elastic body. 
The following will be considered a model problem for a three-dimensional elastic body (also for the two-dimensional elastic body).
\begin{enumerate}
	\item[] The equations of motion \eqref{eq:3D_Model:EM-D} and \eqref{eq:3D_Model:divT-D} is satisfied in $\Omega$;
	\item[] Hooke's law \eqref{eq:3D_Model:CE-D} is satisfied in $\Omega$;
	\item[] The displacement $u = u_\Sigma$ is specified on
		$\Sigma$; The traction $Tn = t_\Gamma$ is specified on
		$\Gamma$.
\end{enumerate} \label{sym:n}\label{sym:SigmaGamma}

\subsubsection{Problem 3D}
Find a vector valued function $u$, satisfying equations
\eqref{eq:3D_Model:EM-D} to \eqref{eq:3D_Model:CE-D} and the following boundary
conditions:\\

Boundary Conditions:
\begin{eqnarray*}
	u & = & 0 \quad \textrm{ on } \Sigma\\
	Tn & = & 0 \quad \textrm{ on } \Gamma
\end{eqnarray*} With $n$ the outward normal vector to $\Omega$. This model problem is a cantilever three-dimensional body.

\subsection{Variational form}\label{ssec:3D_Model:VariationalForm}
Define the bilinear forms
\begin{eqnarray}
	b(u,\phi) = \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dV \label{eq:3D_Model:Bilinear}
\end{eqnarray}\label{sym:b}
and
\begin{eqnarray}
	c(u,\phi) = \int_\Omega~ (\partial^2_t u) \cdot \phi~dV \label{eq:3D_Model:Bilinear_c}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1+\nu)(1-2\nu)}$.\\

Also define the integral
\begin{eqnarray}
	(f,g) &=& \int_{\Omega} f\cdot g \ dV \label{eq:3D_Model:Bilinear_int}
\end{eqnarray}

Define the following set of test functions $T(\Omega)$ for Problem 3D:
\begin{eqnarray*}
	T(\Omega) & = & \left\{ \phi \in C^1(\bar{\Omega}) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}
\end{eqnarray*}\label{sym:TOmega}

The variational problem of Problem 3DV is given as:
\subsubsection{Problem 3DV}\label{sssec:3D_Model:Problem3D1V}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi), \label{eq:3D_Model:Problem3D1VEq}
\end{align}
for all $\phi \in T(\Omega)$\\

\subsection{Plane stress}\label{ssec:3D_Model:PlaneStress}
Stresses acting parallel to a single plane are referred to as plane stress. 
Consider the right-hand orthonormal set $\left\{e_1, e_2, e_3\right\}$. Without
loss of generality, assume the stresses act parallel to the $e_1$-$e_2$ plane
i.e. $\sigma_{3i} = 0$ for all $i$. From Hooke's Law \eqref{DM_H_E} the
following equations are obtained: \label{sym:e_i}

@@

\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\sigma_{11} & = & \frac{1}{\gamma(1-\nu^2)} ( \varepsilon_{11} + \nu\varepsilon_{22})\\
		\sigma_{22} & = & \frac{1}{\gamma(1-\nu^2)} (\nu \varepsilon_{11} + \varepsilon_{22}).
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray}
		\sigma_{12} & = & \frac{1}{\gamma(1+\nu)} \varepsilon_{12} \label{stress_comp}
	\end{eqnarray}
\end{minipage}\\

@@

This is known as general plane stress.

\subsection{Two-dimensional model}
Using \cite{Sadd05} as a guide, a special case of general plane stress is considered. In this special case, a two-dimensional model is obtained.\\

Assume that $\sigma_{3i} = 0$ for $i = 1, 2, 3$; $\partial_3 u_1 = 0$, $\partial_3 u_2 = 0$ ($u_1$ and $u_2$ are functions of $x_1$ and $x_2$), and the strain component $\varepsilon_{33} = 0$. Then using the definition of strain 
\begin{eqnarray}
	\partial_i u_3 = 0 \quad \textrm{ for } i = 1,2,3 \textrm{ on } \Omega.
\end{eqnarray}

@@

\subsection*{Equations of motion and constitutive equations}\label{ssec:2D_Model:EquationsOfMotion+ConstitutiveEquations}
 The following equations of motion and constitutive equations follow from subsections \ref{ssec:3D_Model:DimensionlessForm} and 1.3.2.
\subsubsection{Equation of motion}\label{sssec:2D_Model:EquationOfMotion}
\begin{eqnarray}
	\partial_t^2 u & = & \textrm{div}T + Q, \label{eq:2D_Model:EM}
\end{eqnarray}
where
\begin{eqnarray}
	\textrm{div} T & = &
	\begin{bmatrix}
		\partial_1 \sigma_{11} + \partial_2 \sigma_{12} \\
		\partial_1 \sigma_{21} + \partial_2 \sigma_{22}
	\end{bmatrix}. \label{eq:2D_Model:DivT}
\end{eqnarray}
\subsubsection*{Constitutive equation}\label{sssec:2D_Model:ConstitutiveEquation}
\begin{eqnarray}
	T = \frac{1}{\gamma(1+\nu)}\mathcal{E} + \frac{\nu}{\gamma(1-\nu^2)}\textrm{Tr}(\mathcal{E}) I. \label{eq:2D_Model:CE}
\end{eqnarray}

\subsection{Model problem}\label{ssec:2D_Model:ModelProblem}
Divide the boundary of $\Omega$ into two distinct parts, referred to as $\Sigma$ and $\Gamma$.

\subsubsection*{Problem 2D}\label{sssec:2D_Model:Problem2D1}
Find a vector valued function $u$, satisfying equations \eqref{eq:2D_Model:EM} to \eqref{eq:2D_Model:CE} and the boundary conditions:\\

Boundary Conditions:\\
\begin{eqnarray*}
	u & = & 0 \quad \textrm{ on } \Sigma,\\
	Tn & = & 0 \quad \textrm{ on } \Gamma.
\end{eqnarray*} with $n$ the outward unit normal. This model problem is a cantilever two-dimensional body.

\subsection{Variational form}\label{ssec:2D_Model:VariationalForm}
Define the bilinear forms
\begin{eqnarray}
	b(u,\phi) = \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dA \label{eq:2D_Model:Bilinear}
\end{eqnarray}
and
\begin{eqnarray}
	c(u,\phi) = \int_\Omega~ (\partial^2_t u) \cdot \phi~dA \label{eq:2D_Model:Bilinear_c}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1-\nu^2)}$.\\
Also define the integral
\begin{eqnarray}
	(f,g) &=& \int_{\Omega} f\cdot g \ dA \label{eq:2D_Model:Bilinear_int}
\end{eqnarray}

\subsubsection{Test functions for Problem 2D}\label{sssec:2D_Model:TestFunctions2D1}
Define the following set of test functions $T(\Omega)$ for Problem 2D:
\begin{eqnarray*}
	T(\Omega) & = & \left\{ \phi \in C^1(\bar{\Omega}) \ | \ \phi = 0 \ \textrm{ on } \ \Gamma \right\}. 
\end{eqnarray*}\label{sym:baromega}

The variational problem of Problem 2D is given as:

\subsubsection{Problem 2DV}\label{sssec:2D_Model:Problem2D1V}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi) \label{eq:2D_Model:Problem2D1VEq}
\end{align}
for all $\phi \in T(\Omega)$.\\

\section{Existance and uniqueness of solutions}
Our two models, the three-dimensional modela and the two-dimensional model, concern the vibration of elastic bodies. These type of models have a similar variational form as the wave equation and are called second order hyperbolic type problems. The articles \cite{VV02} and \cite{VS19} provide existence and uniqueness results for these type of problems. In \cite{VV02}, it is assumed that $b$ is symmetric. \cite{VS19} extends the results of \cite{VV02} to the case where $b$ is not symmetric. For the models in the article, \cite{VV02} is a sufficient reference.\\

@@

Consider a general second order hyperbolic type problem GVar. This general problem is used by the 
authors of \cite{VV02} to prove existence and uniqueness results. The models in this article are 
special cases of this general problem.\\
